---
title: "HAR Prediction"
output: html_document
---

The machine learning task at hand is predicting the movement a subject made given data from a number of sensors.  It is a supervised classification task with mostly continuous features.

```{r, echo=FALSE, cache=TRUE}
library(caret)
library(gbm)

setwd("~/Dropbox/RR/Grad school/CS/R")
har <- read.csv("pml-training.csv")
obs <- dim(har)[1]
vars <- dim(har)[2]
har2 <- har[,8:160]
inTrain <- createDataPartition(y=har2$classe, p=0.75, list=FALSE)
training_set <- har2[inTrain,]
crossval_set <- har2[-inTrain,]

t2 <- training_set[,c("roll_belt","pitch_belt","yaw_belt","total_accel_belt","gyros_belt_x","gyros_belt_y","gyros_belt_z","accel_belt_x","accel_belt_y","accel_belt_z","magnet_belt_x","magnet_belt_y","magnet_belt_z","roll_arm","pitch_arm","yaw_arm","total_accel_arm","gyros_arm_x","gyros_arm_y","gyros_arm_z","accel_arm_x","accel_arm_y","accel_arm_z","magnet_arm_x","magnet_arm_y","magnet_arm_z","roll_dumbbell","pitch_dumbbell","yaw_dumbbell","total_accel_dumbbell","gyros_dumbbell_x","gyros_dumbbell_y","gyros_dumbbell_z","accel_dumbbell_x","accel_dumbbell_y","accel_dumbbell_z","magnet_dumbbell_x","magnet_dumbbell_y","magnet_dumbbell_z","roll_forearm","pitch_forearm","yaw_forearm","total_accel_forearm","gyros_forearm_x","gyros_forearm_y","gyros_forearm_z","accel_forearm_x","accel_forearm_y","accel_forearm_z","magnet_forearm_x","magnet_forearm_y","magnet_forearm_z","classe")]

c2 <- crossval_set[,c("roll_belt","pitch_belt","yaw_belt","total_accel_belt","gyros_belt_x","gyros_belt_y","gyros_belt_z","accel_belt_x","accel_belt_y","accel_belt_z","magnet_belt_x","magnet_belt_y","magnet_belt_z","roll_arm","pitch_arm","yaw_arm","total_accel_arm","gyros_arm_x","gyros_arm_y","gyros_arm_z","accel_arm_x","accel_arm_y","accel_arm_z","magnet_arm_x","magnet_arm_y","magnet_arm_z","roll_dumbbell","pitch_dumbbell","yaw_dumbbell","total_accel_dumbbell","gyros_dumbbell_x","gyros_dumbbell_y","gyros_dumbbell_z","accel_dumbbell_x","accel_dumbbell_y","accel_dumbbell_z","magnet_dumbbell_x","magnet_dumbbell_y","magnet_dumbbell_z","roll_forearm","pitch_forearm","yaw_forearm","total_accel_forearm","gyros_forearm_x","gyros_forearm_y","gyros_forearm_z","accel_forearm_x","accel_forearm_y","accel_forearm_z","magnet_forearm_x","magnet_forearm_y","magnet_forearm_z","classe")]

```

# Pre-processing #

The "training"" dataset contains `r obs` observations and `r vars` variables.  We split the training dataset further up into a new true training set (75%) and a cross-validation set (25%).

The first seven variables (columns) do not contain variables that are relevant for (generalized) prediction, as they are e.g. the index number, time stamps, and subject names.

A summary of the dataset and its variables (not shown here for space reasons) indicates that, of the 153 feature variables in the dataset, only 52 do not contain a very significant number of NAs.  We therefore only consider these 52 features for training.

# Training #

## Simple tree ##

As a baseline, we train a simple tree.  We use Caret's confusionMatrix function to output a number of relevant metrics for the evaluation of the trained classifier.

```{r, echo=FALSE, cache=TRUE}
fit_tree <- train(classe~.,data=t2,method="rpart")
preds_tree <- predict(fit_tree,newdata=c2)
confusionMatrix(c2$classe,preds_tree)
```

## Boosted tree ##

We can see from the confusion matrix of the simple tree that the classifier still confuses the classes a lot.  We therefore now consider a boosted tree.  Its output follows below.

```{r, echo=FALSE, cache=TRUE}
fit_boost <- train(classe~.,data=t2,method="gbm")
preds_boost <- predict(fit_boost,newdata=c2,verbose=FALSE)
confusionMatrix(c2$classe,preds_boost)
```

It is evident that the boosted tree improves the classification performance significantly, e.g. raising overall acuracy from 0.50 to 0.97.  This is a good performance by any standard and we will use this classifier on the test set.

# Classification of test set #

The test set predictions follow below.

```{r, echo=FALSE, cache=TRUE}
test_set <- read.csv("pml-testing.csv")
test2 <- test_set[,c("roll_belt","pitch_belt","yaw_belt","total_accel_belt","gyros_belt_x","gyros_belt_y","gyros_belt_z","accel_belt_x","accel_belt_y","accel_belt_z","magnet_belt_x","magnet_belt_y","magnet_belt_z","roll_arm","pitch_arm","yaw_arm","total_accel_arm","gyros_arm_x","gyros_arm_y","gyros_arm_z","accel_arm_x","accel_arm_y","accel_arm_z","magnet_arm_x","magnet_arm_y","magnet_arm_z","roll_dumbbell","pitch_dumbbell","yaw_dumbbell","total_accel_dumbbell","gyros_dumbbell_x","gyros_dumbbell_y","gyros_dumbbell_z","accel_dumbbell_x","accel_dumbbell_y","accel_dumbbell_z","magnet_dumbbell_x","magnet_dumbbell_y","magnet_dumbbell_z","roll_forearm","pitch_forearm","yaw_forearm","total_accel_forearm","gyros_forearm_x","gyros_forearm_y","gyros_forearm_z","accel_forearm_x","accel_forearm_y","accel_forearm_z","magnet_forearm_x","magnet_forearm_y","magnet_forearm_z")]
preds_test <- predict(fit_boost,newdata=test2)
preds_test